# Linear regression

```{r load_libraries, echo = FALSE, warning = FALSE}
library(ggplot2)
library(GGally)
```

```{r load_data, echo = FALSE}
learning2014 <- read.csv("data/learning2014.csv")
```

We are looking at the dataset learning2014. Here is a summary of the dataset:

```{r structural_overview, echo = FALSE}
str(learning2014)
```

This dataset of learning results (as measured by exam points) and possibly explanatory variables consists of 166 observations and 7 variables. Here is a graphical overview of the dataset:

```{r graphical_overview, echo = FALSE}
ggpairs(learning2014, mapping = aes(col = gender, alpha = 0.2), lower = list(combo = wrap("facethist", bins = 25)))
```

There is somewhat high correlation between attitude and points, but otherwise the variables are pretty uncorrelated. All variables have mostly similar distributions for both genders. Genders are marked with coloring.

```{r multiple_regression, echo = FALSE}
first_model <- lm(learning2014$points ~ learning2014$attitude + learning2014$stra + learning2014$surf)
summary(first_model)
```

We first try multiple regression, but strategic and surface learning do not have statistical significance as explainers of exam points. We thus limit ourselves to simple regression with attitude:

```{r simple_regression, echo = FALSE}
second_model <- lm(learning2014$points ~ learning2014$attitude)
summary(second_model)
```

There is very high statistical significance for attitude as explainer of exam points.

```{r fit_plot, echo = FALSE}
plot(second_model, which = 1)
```

Constant variance of errors assumption holds well.

```{r qq_plot, echo = FALSE}
plot(second_model, which = 2)
```

The QQ plot implies that the errors have reasonably normal distribution, though there are some problems in both extremes.

```{r leverage_plot, echo = FALSE}
plot(second_model, which = 5)
```

No observation has abnormally large leverage, making the model seem more reasonable.
